---
title: "STA 141A - Spring 2024 - Final Project"
date: 'Due date: Apr 10, 2024 at 07:59 PM (PT)'
output:
  html_document:
    df_print: paged
subtitle: 'Understanding Factors Influencing College Student Performance: 
  Insights and Recommendations for Students Success Initiatives'
---

## 2.1 Correlation between prior academic Performance and college marks (Lola)

```{r}
# Data Import and Structure
library(ggplot2)
library(dplyr)
library(gridExtra)

filename <- "Student Attitude and Behavior.csv"
file_path <- file.choose()
if (basename(file_path) == filename) {
  data <- read.csv(file_path, header = TRUE)
} else {
  cat("Selected file does not match expected filename.\n")
}
head(data)

str(data)

# Data Transform
names(data) <- c('Certification_Course','Gender','Department','Height','Weight',
                 'Mark_10','Mark_12','College_Mark', 'Hobbies', 'Daily_Study_Time', 
                 'Study_Spot', 'Salary_Expectation', 'Degree_Opinion', 'Opinion_Future_Career', 
                 'Social_Media_Engagement', 'Travel_Time', 'Stress_Level', 'Financial_Status', 
                 'Job')

```

### 2.1.1 College Marks vs Prior Academic Performance (Lola)
```{r}
# Scatter Plots
par(mfrow = c(1, 2))

plot(data$Mark_10, data$College_Mark, 
     main = "College Mark vs 10th Grade Mark", ylab = "College Mark",
     xlab = "10th Grade Mark", pch = 19, font = 1, font.lab = 2,
     cex = 0.75)
plot(data$Mark_12, data$College_Mark, 
     main = "College Mark vs 12th Grade Mark", ylab = "College Mark",
     xlab = "12th Grade Mark", pch = 19, font = 1, font.lab = 2,
     cex = 0.75)

par(mfrow = c(1, 1))

# Regression Models
the.model_1.1 <- lm(College_Mark ~ Mark_10, data = data)
the.model_1.2 <- lm(College_Mark ~ Mark_12, data = data)
summary(the.model_1.1)
summary(the.model_1.2)
```


## 2.2 Correlation between personal factors and college marks
```{r}
## exploratory data visualization
par(mfrow = c(2, 2))
## Do you like your degree?
plot1 = ggplot(data, aes(x = factor(Degree_Opinion))) +
  geom_bar(fill = "bisque", color = "black") +
  labs(title = "Do you like your degree?",
       x = "Do you like your degree?",
       y = "Count")
##  Willingness to pursue a career based on their degree
plot2 = ggplot(data, aes(x = factor(Opinion_Future_Career))) +
  geom_bar(fill = "deeppink", color = "black") +
  labs(title = "Would You Pursue a Career Based on Your Degree?",
       x = "Willingness",
       y = "Count")+
  theme(plot.title = element_text(size = 9.5))
## Salary Expectations
breaks <- seq(0, 150000, by = 10000)  # Adjust the range and width of bins as needed
# Create the histogram
plot3= ggplot(data, aes(x = Salary_Expectation)) +
  geom_histogram(bins = length(breaks) - 1, fill = "cadetblue", color = "black", breaks = breaks) +
  labs(title = "Histogram of Salary Expectation",
       x = "Salary Expectation",
       y = "Count")
## Stress Level
plot4 = ggplot(data, aes(x = factor(Stress_Level))) +
  geom_bar(fill = "pink", color = "black") +
  labs(title = "Stress Level",
       x = "Stress Level",
       y = "Count")

#arrange plots       
grid.arrange(plot1, plot2, plot3, plot4, nrow = 2, ncol = 2)
```

### 2.2.1 College Marks vs Salary Expectation (Aidan)
```{r}
library(dplyr)
library(car)

#Check for missing values and omit NAs
missing_values <- colSums(is.na(data))
data_no_NA <- na.omit(data)

#Check to make sure there are indeed no missing values
missing_values_check <- colSums(is.na(data_no_NA))
missing_values_check

# Table 2.1
College_Mark <- data$College_Mark
Salary_Expectation <- data$Salary_Expectation
data1 <- data.frame(
  college_marks = College_Mark,
  salary_expectation = Salary_Expectation
)

data1 <- data1 %>%
  mutate(College_Marks_Category = cut(college_marks, 
                              breaks = c(-Inf, 80, 90, Inf), 
                              labels = c("Low", "Medium", "High")))

Average_Salary <- data1 %>%
  group_by(College_Marks_Category) %>%
  summarise(Avg_Salary_Expectation = mean(salary_expectation))

print(Average_Salary)

library(knitr)
kable(Average_Salary, caption = "Table 2.3.1: Average Salary Expectations by College Marks Category")

# Plot 2.1
library(ggplot2)

ggplot(data = Average_Salary, aes(x = College_Marks_Category, y = Avg_Salary_Expectation, fill = College_Marks_Category)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(Avg_Salary_Expectation, 2)), position = position_stack(vjust = 0.5), color = "black", size = 3) +
  labs(title = "Average Expected Salary by College Marks Category",
       x = "College Marks Category",
       y = "Average Expected Salary") +
  scale_fill_manual(values = c("Low" = "red", "Medium" = "blue", "High" = "green")) +
  theme(legend.position = "none")

# Linear Regression Plot
lm_model <- lm(salary_expectation ~ college_marks, data = data1)

summary(lm_model)

ggplot(data1, aes(x = college_marks, y = salary_expectation)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(x = "College Marks", y = "Salary Expectation") + 
  ggtitle("Linear Regression: College Marks vs. Salary Expectation") +
  theme_minimal() 

```

### 2.2.2 College Marks vs Willingness to pursue a career based on their degree
```{r}
College_Mark <- data$College_Mark
Willingness <- data$Opinion_Future_Career
willingness_numeric <- as.numeric(gsub("%", "", Willingness))
data2 <- data.frame(
  college_marks = College_Mark,
  willingness = willingness_numeric
)
library(ggplot2)
library(dplyr)

model <- lm(willingness ~ college_marks, data = data2)
summary(model)

ggplot(data2, aes(x = college_marks, y = willingness)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "College Marks vs. Willingness to Pursue a Career Based on Their Degree",
    x = "College Marks",
    y = "Willingness (%)"
  ) +
  theme_minimal()

plot(model)
qqnorm(residuals(model))
qqline(residuals(model))
```

### 2.2.3 College Marks v.s. Stress Level (Emma)
```{r}
# Table 2.2
library(knitr)
mark_by_stress = data %>% 
                    group_by(Stress_Level) %>%
                    summarize(AvgMark10 = mean(Mark_10),
                             AvgMark12 = mean(Mark_12),
                             AvgCollege_Mark = mean(College_Mark))
kable(mark_by_stress, format = "markdown")
mark_by_stress

# Plot 2.2
##VISUALIZATION
# average 10th grade mark by stress level plot 
plot_10 = ggplot(mark_by_stress) + 
  geom_col(aes(x = Stress_Level , y = AvgMark10, fill = Stress_Level)) +
  geom_text(aes(x = Stress_Level , y = AvgMark10, label = round(AvgMark10, 2)), vjust = 0.5, size = 3) +
  scale_fill_manual(values=c("#99CCFF", 
                             "#CCCFFF", 
                             "#FF99CC", 
                             "#FFCCFF")) +
  theme_minimal() + 
  labs(title = 'Average 10th Grade Mark by Stress Level',
       x = "Stress Level",
       y = "Average 10thth Grade Mark") +
  theme(title = element_text(size =9),
    axis.title = element_text(size = 8), 
    axis.text.x = element_text(hjust = 1, size =5),
    legend.position = "none")

# average 12th grade mark by stress level plot 
plot_12 = ggplot(mark_by_stress) + 
  geom_col(aes(x = Stress_Level , y = AvgMark12, fill = Stress_Level)) +
  scale_fill_manual(values=c("#99CCFF", 
                             "#CCCFFF", 
                             "#FF99CC", 
                             "#FFCCFF")) +
  geom_text(aes(x = Stress_Level , y = AvgMark12, label = round(AvgMark12, 2)), vjust = 0.5, size = 3) +
  theme_minimal() + 
  labs(title = 'Average 12th Grade Mark by Stress Level',
       x = "Stress Level",
       y = "Average 12th Grade Mark") +
  theme(title = element_text(size =9),
        axis.title = element_text(size = 8), 
        axis.text.x = element_text(size = 5),
        legend.position = "none")

# average college mark by stress level plot 
college_plot = ggplot(mark_by_stress) + 
  geom_col(aes(x = Stress_Level , y = AvgCollege_Mark, fill = Stress_Level)) +
  scale_fill_manual(values=c("#99CCFF", 
                             "#CCCFFF", 
                             "#FF99CC", 
                             "#FFCCFF")) +
  geom_text(aes(x = Stress_Level , y = AvgCollege_Mark, label = round(AvgCollege_Mark, 2)), vjust = 0.5, size = 3) +
  theme_minimal() + 
  labs(title = 'Average College Mark by Stress Level',
       x = "Stress Level",
       y = "Average College Mark") +
  theme(title = element_text(size =9),
        axis.title = element_text(size = 8), 
        axis.text.x = element_text(size = 5),
        legend.position = "none")

# arrange the plots
grid.arrange(plot_10, plot_12, college_plot, ncol = 2)

# Logistic Regression 
## To convert stress into binary variable: 
data$Quantified_Stress = ifelse(data$Stress_Level %in% c("fabulous", "Good"), 0, 1)
## logistic regression
college_logreg = glm(Quantified_Stress ~College_Mark, data = data, family = binomial)
## Summary of the logistic regression model
summary(college_logreg)

# Plot 2.3
ggplot(data, aes(x=College_Mark, y=Quantified_Stress)) + geom_point() +
      stat_smooth(method="glm", color="blue", se=FALSE, 
                method.args = list(family=binomial))

```

### 2.2.4 Stress Levels v.s. Personal Factors (Emma)
```{r}
# Plot 2.4
library(ggplot2)
library(dplyr)
# How does having a part time job affect stress levels?
ggplot(data, aes(x = factor(Job), fill = factor(Stress_Level))) +
  geom_bar(position = "dodge") +
  labs(title = "Stress Level by Part-time Job",
       x = "Part-time Job?",
       y = "Count") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

## How does financial status affect stress levels?
ggplot(data, aes(x = factor(Financial_Status), fill = factor(Stress_Level))) +
  geom_bar(position = "dodge") +
  labs(title = "Stress Level by Financial Status",
       x = "Financial Status",
       y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# How does travel time/commute to educational institutions affect stress levels?
ggplot(data, aes(x = factor(Travel_Time), fill = factor(Stress_Level))) +
  geom_bar(position = "dodge") +
  labs(title = "Stress Level by Travel Time to Educational Institution",
       x = "Commute Time",
       y = "Count") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Logistic Regression #2 
## To quantify stress levels, convert into binary variable 
data$Quantified_Stress = ifelse(data$Stress_Level %in% c("fabulous", "Good"), 0, 1)
## logistic regression
logreg = glm(Quantified_Stress ~ Job + Financial_Status + Travel_Time, data = data, family = binomial)
# Summary of the logistic regression model
summary(logreg)
## Predict probabilities
prob <- predict(logreg, type = "response")
## Assign predicted quantified stress levels
predicted <- ifelse(prob < 0.5, "Low Stress", "High Stress")

## confusion matrix
predicted = predict(logreg, data, type = 'response')
predicted_class = ifelse(predicted>=0.5, 1, 0)
confusion_matrix = table(actual = data$Quantified_Stress, predicted = predicted_class)
confusion_matrix
## calculate accuracy rate 
accuracy = sum(diag(confusion_matrix))/sum(confusion_matrix)
accuracy

```

## 2.3 Correlation between student’s engagement with society and college marks

### 2.3.1 College Marks vs Hobbies (Merry)
```{r}
hobby_counts <- table(data$Hobbies)
hobby_df <- as.data.frame(hobby_counts)

# Bar Plot
barplot(hobby_df$Freq, 
        names.arg = hobby_df$Var1, 
        main = "College Marks vs Hobbies", 
        xlab = "Hobbies", 
        ylab = "College Marks", 
        col = "skyblue")

unique(data$Hobbies)

data <- data %>%
  mutate(B = case_when(
    Hobbies == 'Cinema' ~ 1,
    Hobbies == 'Reading books' ~ 2,
    Hobbies == 'Sports' ~ 3,
    Hobbies == 'Video Games' ~ 4,
    TRUE ~ NA_integer_
  ))
data$B <- as.numeric(data$B)

# Scatter Plot
plot(data$B,data$College_Mark,main = "College Mark vs Hobbies", 
     ylab = "College Mark",xlab = "Hobbies")

#Regression
the.model <- lm(College_Mark ~ Hobbies, data = data)
summary(the.model)
```


### 2.3.2 College Marks vs Engagement with Social Media (Lola)
```{r}
# Data Transform
data <- data %>%
  mutate(Social = case_when(
    Social_Media_Engagement == '0 Minute' ~ '0',
    Social_Media_Engagement == '1 - 30 Minute' ~ '15',
    Social_Media_Engagement == '30 - 60 Minute' ~ '30',
    Social_Media_Engagement == '1 - 1.30 hour' ~ '60',
    Social_Media_Engagement == '1.30 - 2 hour' ~ '90',
    Social_Media_Engagement == 'More than 2 hour' ~ '120'
  )) %>%
  select(-Social_Media_Engagement)
data$Social <- as.numeric(data$Social)

unique_values2 <- unique(data$Social)
social_counts <- table(data$Social)
social_time_df <- as.data.frame(social_counts)
social_time_df$Var1 <- factor(social_time_df$Var1, levels = unique(social_time_df$Var1))

# Scatter Plot
plot(data$Social, data$College_Mark, 
     main = "College Mark vs Entertainment Time", ylab = "College Mark", 
     xlab = "Entertainment Time", pch = 19, font = 2, font.lab = 2, cex = 0.75)
# Bar Plot
barplot(social_time_df$Freq, 
        names.arg = social_time_df$Var1, 
        main = "Histogram of Entertainment Time", 
        xlab = "Entertainment Time", 
        ylab = "Frequency")
# Regression
the.model_2.1 <- lm(College_Mark ~ Social, data = data)
summary(the.model_2.1)
```

